---
title: "marketing_data_eda"
author: "Kevin"
date: "30 May 2021"
output: rmarkdown::github_document
---

```{r, warning=FALSE}
library(tidyverse)
library(data.table)
library(lubridate)
library(corrplot)
```

```{r}
marketing_raw <- fread('data/marketing_data.csv', colClasses = c('Education'='factor', 'Marital_Status'='factor',
                                                                 'Country'='factor'))
head(marketing_raw)
```

## About data - 
Data is from a iFood CRM(mentioned in kaggle discussions). It is a Brazilian food delivery platform. Data is of customers who made purchase in the last 100 days.

## Classes of variables - 
1) Customer Attributes
2) Amount spent on different product groups
3) Number of Purchases through Channels and visits
4) Success after each campaign

The means of AcceptedCmp and Response and Complain will give the percentage of 1's in the columns.
Most AcceptedCmp have a 7% success rate. 

```{r}
marketing_proc <- marketing_raw %>% 
  mutate(Dt_Customer = mdy(Dt_Customer),  
         Income = as.double(str_replace_all(Income, "[$,]", "")),
         Age=2014-Year_Birth,
         People_at_home = if_else(Marital_Status %in% c('Married','Together'), Kidhome+Teenhome+2, Kidhome+Teenhome+1),
         Age_bin = case_when(Age<35 ~ '<35',
                             Age<45 ~ '35-44',
                             Age<60 ~ '45-59',
                             TRUE ~ '>=60'),
         Income_bin = case_when(Income<35000 ~ '<35k',
                                Income<65000 ~ '35k-65k',
                                TRUE ~ '>=65k'),
         Response=case_when((Response==1 | AcceptedCmp1==1 | AcceptedCmp2==1 | 
                              AcceptedCmp3==1 | AcceptedCmp4==1 | AcceptedCmp5==1) ~ 1,
                            TRUE ~ 0)) 
# Assuming data is extracted around 30th June 2014
# Assuming parents of customers do not stay with them
summary(marketing_proc)
```

### Response is 1 if any of the accepted or response variables is 1. Thus, response rate is mean of that column, 27%

## Customer Attributes

### Age

```{r}
qplot(marketing_proc$Age, geom='histogram', xlim=c(15,80), bins=20)
```

Biggest customer group is in age group of 35 to 45. Few people above 100 yrs old are excluded from the view for better view.

### Education

```{r}
qplot(marketing_proc$Education, geom="bar")
```

Customer base seems to be atleast Graduates mostly

### Income

```{r}
qplot(marketing_proc$Income, geom='histogram', xlim=c(0,200000))
```

Income is nearly symmetrically distributed around \$50000 with a few people around \$150,000 and one person with over \$600,000

### Marital Status

```{r}
qplot(marketing_proc$Marital_Status, geom="bar")
```

Mostly Married and Together and Divorced, so consistent with the middle aged group

### People at Home

```{r}
qplot(marketing_proc$People_at_home, geom='bar')
```

Mostly 2 or 3 people in a family(Mostly one person would make food purchases for entire family)

### Country

```{r}
qplot(marketing_proc$Country, geom='bar')
```

Over half the customer base is from SP(possibly Spain)

## Purchases

### Amount spent on Fruits

```{r}
qplot(marketing_proc$MntFruits, geom = 'histogram')
```

Mostly between \$0 to \$25

### Amount spent on Fish

```{r}
qplot(marketing_proc$MntFishProducts, geom = 'histogram')
```

Mostly between \$0 to \$40

### Amount spent on Meat

```{r}
qplot(marketing_proc$MntMeatProducts, geom = 'histogram')
```

Mostly between \$0 to \$25 but people may spend upto \$1500 on meat

### Amount spent on Sweets

```{r}
qplot(marketing_proc$MntSweetProducts, geom = 'histogram')
```

Mostly between \$0 to \$50

### Amount spent on Wines

```{r}
qplot(marketing_proc$MntWines, geom = 'histogram')
```

Spending on Wine is huge, even upto $1000 by some

### Amount spent on Gold

```{r}
qplot(marketing_proc$MntGoldProds, geom = 'histogram')
```

People are mostly spending upto $50 on gold

## Which demographic is most likely to give a response?

```{r}
marketing_proc %>% 
  group_by(Age_bin) %>%
  summarise(avg_response = mean(Response)) %>%
  ggplot() +
  geom_col(aes(x=Age_bin, y=avg_response)) 
```

```{r}
marketing_proc %>% 
  group_by(Income_bin) %>%
  summarise(avg_response = mean(Response)) %>%
  ggplot() +
  geom_col(aes(x=Income_bin, y=avg_response)) 
```

Response is significantly higher in the highest income group

```{r}
marketing_proc %>% 
  group_by(People_at_home) %>%
  summarise(avg_response = mean(Response)) %>%
  ggplot() +
  geom_col(aes(x=People_at_home, y=avg_response)) 
```

Highest response rate is in the below 35 age group.

```{r}
marketing_proc %>% 
  group_by(Age_bin, Income_bin) %>%
  summarise(avg_response = mean(Response)) %>%
  ggplot() +
  geom_col(aes(x=Age_bin, y=avg_response, fill=Income_bin), position='dodge') 
```

```{r}
marketing_proc %>% 
  group_by(People_at_home, Income_bin) %>%
  summarise(avg_response = mean(Response)) %>%
  ggplot() +
  geom_col(aes(x=People_at_home, y=avg_response, fill=Income_bin), position='dodge') 
```

## Correlation
```{r, fig.width = 14, fig.height = 11}
fil <- marketing_proc %>% select(-c(Country, Marital_Status, Education, Dt_Customer, Age_bin, Income_bin)) %>%
  filter(!is.na(Income))
M <- cor(fil)
corrplot(M)
```

## Who's making the most purchase in each Product category?(So that we can market the right product to the right demographic)

```{r}
marketing_proc %>% 
  group_by(Age_bin) %>%
  summarise(avg_fruit_expense=mean(MntFruits),
            avg_fish_expense=mean(MntFishProducts),
            avg_meat_expense=mean(MntMeatProducts),
            avg_wine_expense=mean(MntWines),
            avg_gold_expense=mean(MntGoldProds),
            avg_sweet_expense=mean(MntSweetProducts)) %>%
  gather("Group","Expense", -Age_bin) %>%
  ggplot() +
  geom_col(aes(x=Age_bin, y=Expense, fill=Group), position = 'dodge') 
```

Older People spend more on Wine

```{r}
marketing_proc %>% 
  group_by(Income_bin) %>%
  summarise(avg_fruit_expense=mean(MntFruits),
            avg_fish_expense=mean(MntFishProducts),
            avg_meat_expense=mean(MntMeatProducts),
            avg_wine_expense=mean(MntWines),
            avg_gold_expense=mean(MntGoldProds),
            avg_sweet_expense=mean(MntSweetProducts)) %>%
  gather("Group","Expense", -Income_bin) %>%
  ggplot() +
  geom_col(aes(x=Income_bin, y=Expense, fill=Group), position = 'dodge') 
```

Richer people spend more on all product groups but especially meat and wine

## Who uses which channel more?
```{r}
marketing_proc %>% 
  group_by(People_at_home) %>%
  summarise(avg_deals_purchases=mean(NumDealsPurchases),
            avg_web_purchases=mean(NumWebPurchases),
            avg_catalog_purchases=mean(NumCatalogPurchases),
            avg_store_purchases=mean(NumStorePurchases)) %>%
  gather("Group","Purchases", -People_at_home) %>%
  ggplot() +
  geom_col(aes(x=People_at_home, y=Purchases, fill=Group), position = 'dodge') 
```

When there are more people in the house, people prefer deals over catalogs.

```{r}
marketing_proc %>% 
  group_by(Income_bin) %>%
  summarise(avg_deals_purchases=mean(NumDealsPurchases),
            avg_web_purchases=mean(NumWebPurchases),
            avg_catalog_purchases=mean(NumCatalogPurchases),
            avg_store_purchases=mean(NumStorePurchases),
            avg_web_visits=mean(NumWebVisitsMonth)) %>%
  gather("Group","Purchases", -Income_bin) %>%
  ggplot() +
  geom_col(aes(x=Income_bin, y=Purchases, fill=Group), position = 'dodge') 
```

People with higher income spend more but people with lower income visit the site more.

